---
title: "Synechocystis_Data_Analysis"
author: "T.A.J. Uijen, "
date: "01/04/2025-29/05/2026"
output:
  pdf_document:
    number_sections: no
    toc: yes
    toc_depth: 3
  html_document:
    toc: yes
    toc_depth: '3'
    df_print: paged
editor_options:
  markdown:
    wrap: 72
execute: 
  echo: true
  message: false
  warning: false
---

# Importing data and loading packages

### Packages

```{r, echo=TRUE, message=FALSE}
library(readr)
library(tidyr)
library(dplyr)
library(ggplot2)
library(flextable)
library(GGally)
library(broom)
library(readxl)
```

### Data

```{r}
Design<-read_csv("Data/Design.csv")
SpecPhot<-read_csv("Data/Spectrophotometer.csv")
Turbidostat<-read_csv("Data/Turbidostat.csv")
CASY<-read_csv("Data/Casy.csv")
qPCR<-read_csv("Data/Qpcr.csv")
```

# Analysis of turbidostat and OD data

## Spectrophotometer data

### A bit of data wrangling

```{r}
Design<-Design|>
  mutate(phosphate_mM=factor(phosphate_mM)) # Setting factor so the variable is discrete

# Compute the chlorophil A and phycobilisomes concentrations
SpecPhot_total<-SpecPhot|>
  arrange(channel)|>
  mutate(phosphate_mM=Design$phosphate_mM,
         chlor_a=10.186*(od_value_685-od_value_730)-0.08,
         PBS=(od_value_634-od_value_730)/od_value_730)
  
# Compute the mean for each treatment group
SpecPhot_treatment<-SpecPhot_total|>
  group_by(phosphate_mM)|>
  summarise(od_value_730=mean(od_value_730),
            od_value_685=mean(od_value_685),
            od_value_634=mean(od_value_634),
            od_value_720=mean(od_value_720),
            chlor_a=mean(chlor_a),
            PBS=mean(PBS))


```

### Visualising the data

```{r}
# Make the plots 
od_730_plot<-ggplot(SpecPhot_treatment, aes(phosphate_mM,od_value_730,fill = phosphate_mM))+
  geom_col(show.legend = F)+
  labs(title="Bar plot",
       subtitle = expression(OD[730]~"per treatment group"),
       x="[Phosphate] (mM)", 
       y=expression(OD[730]))+
  scale_fill_brewer(type="seq",palette = "Dark2") +
  theme_bw()


chlor_a_plot<-ggplot(SpecPhot_treatment, 
                     aes(phosphate_mM,chlor_a,fill = phosphate_mM))+
  geom_col(show.legend = F)+
  labs(x="[Phosphate] (mM)", 
       y=expression("Chlorophyl A (mg/L)"),
       title="Bar plot",
       subtitle = "Chorophyl A per treatment group")+
  scale_fill_brewer(type="seq",palette = "Dark2")+
  theme_bw()

pbs_plot<-ggplot(SpecPhot_treatment, 
                 aes(phosphate_mM,PBS,fill = phosphate_mM))+
  geom_col(show.legend = F)+
  labs(x="[Phosphate] (mM)", 
       y=expression("Phycobilisomes (mg/L)"),
       title="Bar plot",
       subtitle = "Phycobilisomes per treatment group")+
  scale_fill_brewer(type="seq",palette = "Dark2")+
  theme_bw()

od_730_plot
chlor_a_plot
pbs_plot
```

Observations:

-   The OD is higher in the group with more phosphate

-   The Chlorophil A content is higher in the group with more phosphate.
    This is noteworthy since there is less light provided to the
    cyanobacteria here. One might propose that the bacteria attempt to
    get more out of the limited light?

-   Phycobilisome content does not change significantly, this is also
    shown with a t-test below

### Hypothesis tests

```{r}
test_OD<-t.test(od_value_730~phosphate_mM,SpecPhot_total)
test_chlorA<-t.test(chlor_a~phosphate_mM,SpecPhot_total)
test_PBS<-t.test(PBS~phosphate_mM,SpecPhot_total)

spec_phot_test_results<-data.frame(Variable=c('OD_730',"Chlorophil A", "Phycobilisomes"), P_value=c(round(test_OD$p.value,3),  round(test_chlorA$p.value,3), round(test_PBS$p.value,3)))

spec_phot_test_results|>flextable()
```

## Turbidostat data

### Initial visualisation of time series

```{r}
Turbidostat<-Turbidostat|>
  full_join(Design|>select(channel,phosphate_mM,purpose))



last_points <- Turbidostat |>
  group_by(channel, decision) |>
  slice_max(time_h, n = 1, with_ties = FALSE)

ggplot(Turbidostat,aes(time_h,log10(OD),colour=phosphate_mM))+
  geom_point()+
  facet_wrap(vars(channel),scales="free_x")+geom_smooth(
  method = "lm",
  aes(group = decision),
  colour = "red",
  alpha = 0.4,
  linewidth = 0.5
)+geom_point(
    data = last_points,
    colour = "cyan",
    size = 1
  ) +
  labs(title="Time series",subtitle=expression(log[10](OD[720]) ~" over time in hours, coloured by treatment group"), x="Time (h)",y=expression(log[10](OD[720])), colour="[phosphate] (mM)")+
  scale_colour_brewer(type="seq",palette = "Dark2")+
  theme_bw()
  
```

Observations:\

-   Channel 5 starts low, behaves normally after the first activation of
    the pump

-   Channel 3 seems translated downwards, as the pump is supposed to
    turn on at OD=0.35, here this seems to happen at OD=0.32. In
    principle this does not change the analysis, as we are concerned
    with the rate of change, not the absolute values. Furthermore, we
    expect a linear relationship between $\text{OD}_{720}$ and cell
    count for measurements between 0.1 and 0.7. Channel 3 falls within
    this range.

-   

### Some data wrangling again

```{r}
Growth_rates<-Turbidostat|>
  mutate(log_OD=log10(OD))|> # compute log10 OD
  group_by(channel,decision)|>
  #Compute growth rates and R^2             
  summarise(mu=coef(lm(log_OD~time_h))[2],
            r_sq=summary(lm(log_OD~time_h))$r.squared, 
            .groups = "drop_last")|>  
  filter(r_sq>.9) |># Remove cycles with less than 0.9 R^2|. This causes ch. 1 to only have 1 point
  left_join(Design |> select(channel, phosphate_mM))


test<-shapiro.test(Growth_rates|>filter(phosphate_mM=="0.0115")|>pull(mu))

Mean_growth<-Growth_rates|>
  group_by(channel)|>
  summarise(mean_mu=mean(mu),
            sd_mu=sd(mu))|>
  left_join(Design |> select(channel, phosphate_mM), by = "channel")|>
  mutate(channel=factor(channel))

  

Growth_rates<-Growth_rates|>
  mutate(channel=factor(channel))

Mean_growth_per_group<-Growth_rates|>
  group_by(phosphate_mM)|>
  summarise(mean_mu_grp=mean(mu),sd_mu_grp=sd(mu),
            CI1=t.test(mu)$conf.int[1],
            CI2=t.test(mu)$conf.int[2]) #Confidence interval is assuming normal distribution of the mean
 





```

### Visualisation

```{r}
violin_plot<-ggplot(Growth_rates,aes(channel,mu,fill=phosphate_mM))+
  geom_violin()+
  geom_point()+
  labs(title="Violin plot",
       subtitle = "Growth rates per channel, coloured by treatment group",
       x="Channel",
       y=expression(mu),
       fill="[phosphate] (mM)")+
  scale_fill_brewer(type="seq",palette = "Dark2")+
  theme_bw()

technical_var<-ggplot(Mean_growth,aes(channel,mean_mu,fill=phosphate_mM))+
  geom_col()+
  geom_errorbar(aes(ymin=mean_mu-2*sd_mu,ymax=mean_mu+2*sd_mu),width=0.2)+
  scale_fill_brewer(type="seq",palette = "Dark2")+
  labs(title="Technical ",
       subtitle = "Mean growth rates per channel, error bars represent 2 standard deviations",
       x="Channel",
       y=expression("mean" ~ mu),
       fill="[phosphate] (mM)")+
  theme_bw()

biological_var<-ggplot(Mean_growth_per_group,aes(phosphate_mM,mean_mu_grp,fill=phosphate_mM))+geom_col(show.legend = F)+geom_errorbar(aes(ymin=CI1,ymax=CI2),width=0.2)+
  scale_fill_brewer(type="seq",palette = "Dark2")+
  labs(title="Bar plot",
       subtitle = "Growth rates per treatment group, error bars represent 95% confidence interval",
       x="[phosphate] (mM)",
       y=expression("mean"~ mu))+
  theme_bw()

violin_plot

technical_var

biological_var

t.test(mu~phosphate_mM,Growth_rates) #the null hypothesis that the mean growthrate between groups is the same is rejected
```

Observations:

-   No measurement of SD of channel 1

-   The other channels seem to not differ significantly (assuming normal
    distributions)

-   The data does not suggest a significant group difference between
    phosphate concentration (this is as desired) This is corroborated by
    a t-test (p-value\>0.05) (under the assumption of normality).

We test the normality assumption using a Shapiro-Wilk normality test,
given the small sample size of 20 growth rates that have an $R^2>0.90$

```{r}
normality_test<-Growth_rates|>
  group_by(phosphate_mM)|>
  summarise(p.value=shapiro.test(mu)$p.value)
```

The test does not suggest non-normality of the data, which speaks in
favour of the reliability of the analysis above.

Since the tests do not suggest non-normality, nor a difference in mean
OD$_{720}$, we can test normality further by not grouping by treatment:

```{r}
shapiro.test(Growth_rates$mu)
```

In this case we also do not reject the null hypothesis of normality.

# Ploidy determination

## CASY-counter data

Assumptions: Cells are spherical, no contaminants not captured in the
blank.

Counts measured by the CASY-counter are in a sample that is undiluted
for the MC samples, 1:50 for the BBB samples and 1:2 for the ABB
samples. 10 $\mu$L of this dilution is then added to 10ml of CASY
buffer. Of this mixture, 200 $\mu$L is measured by the CASY-counter.

First compute the mean cell counts for each channel over the 3
replicates

```{r}
MC_CASY_cellcount<-CASY|>
  filter(sample_type=="MC")|>
  group_by(channel)|>
  summarise(mean_count_ml=mean(1001*5*counts))|> # Multiply by 5 since data comes from 200 uL samples
  left_join(Design)|>
  mutate(channel=factor(channel)) # makes the plots prettier

BBB_cellcount<-CASY|>
  filter(sample_type == "BBB")|>
  group_by(channel)|>
  summarise(BBB_count_ml=mean(1001*5*50*counts))|> # Multiply by 50 to account for dilution, 5 and 1001 same as above
  left_join(Design) 

ABB_BBB_cellcount<-CASY|>
  filter(sample_type=="ABB")|>
  group_by(channel)|>
  summarise(ABB_count_ml=mean(1001*5*2*counts))|> # Multiply by 2 to account for dilution, 5 and 1001 same as above
  left_join(BBB_cellcount) 

Broken_cell_count=ABB_BBB_cellcount|>
  mutate(Broken_cell_count_ml=BBB_count_ml-ABB_count_ml,
         channel=factor(channel),
         lysis_efficiency = 100*Broken_cell_count_ml/BBB_count_ml)
  
ABB_BBB_cellcount<-ABB_BBB_cellcount|>
  pivot_longer(cols = c(ABB_count_ml,BBB_count_ml), # Manipulation for plotting reasons
               names_to = "Bead_status",
               values_to = "count_ml") |>
  mutate(channel=factor(channel),
         Bead_status=factor(Bead_status))
 
  



```

```{r}
MC_cellcount_plot<-ggplot(MC_CASY_cellcount,aes(channel,mean_count_ml,fill=phosphate_mM))+geom_col()+
  scale_fill_brewer(type="seq",palette = "Dark2",name = "[Phosphate] (mM)") +
  theme_bw()+
  labs(title="Cell counts in the MultiCultivator per channel",
       subtitle = "Computed from CASY-counts",
       x="Channel",
       y= "Cell concentration (cells/ml)")


comp_cellcount_plot<-ggplot(ABB_BBB_cellcount,aes(channel,count_ml,fill=Bead_status))+
  geom_col(position = "dodge") +
  scale_fill_brewer(type="seq",palette = "Set1",name=" Bead beating", labels=c('ABB', "BBB")) +
  labs(title="Cell counts before and after bead beating per channel",
       subtitle = "Computed from CASY-counts",
       x="Channel",
       y= "Cell concentration (cells/ml)") + 
  theme_bw()

Broken_cell_count_plot<-ggplot(Broken_cell_count,aes(channel,Broken_cell_count_ml,fill=phosphate_mM))+geom_col()+
  scale_fill_brewer(type="seq",palette = "Dark2",name = "[Phosphate] (mM)") +
  labs(title="Broken cell counts per channel",
       subtitle = "Computed from CASY-counts",
       x="Channel",
       y= "Cell concentration (cells/ml)") +
  theme_bw()

lysis_efficiency_plot<-ggplot(Broken_cell_count,aes(channel,lysis_efficiency,fill=phosphate_mM))+geom_col()+
  scale_fill_brewer(type="seq",palette = "Dark2",name = "[Phosphate] (mM)") +
  labs(title="Lysis efficiency per channel",
       subtitle = "Computed from CASY-counts",
       x="Channel",
       y= "Lysis efficiency (%)") +
  theme_bw()

MC_cellcount_plot

comp_cellcount_plot

Broken_cell_count_plot
lysis_efficiency_plot
```

## qPCR Data:

The molecular weight of the DNA was computed using the tool at
<https://oligocalc.eu/>, with the settings set to dsDNA, no primers
(since the primers are in the mastermix) nor modifications at either '
end (since those are also in the mastermix).

We begin by computing the standard curves for each replicate, and the
relevant metrics to determine which curve to use.

### Computation

```{r}
MolW_DNA<-670331.8 # Computed using oligonucleotide tool at https://oligocalc.eu/
Na=6.02214076*10^23 # Avogadro's number

# Compute the dna concentration in mg per uL
qPCR_standard<-qPCR|>
  filter(is.na(channel))|>
  mutate(dna_ng=0.04/(dilution_factor*5), # Factor of 5 due to the 1:5 mixture with the master mix
         dna_fragment=dna_ng/(10^9)/MolW_DNA*Na, # Compute 
         replicate=factor(replicate))
  

# Add copy of all rows with replicate "all" for plotting purposes
qPCR_standard <- bind_rows(qPCR_standard, qPCR_standard |>
  mutate(replicate = "all",
         replicate = factor(replicate)))


qPCR_standard_metrics<-qPCR_standard|>
  group_by(replicate)|>
  summarise(slope = coef(lm(Ct ~ log10(dna_fragment)))[2],
            E= (10^(-1/slope)/2)*100,
            R2 = summary(lm(Ct ~ log10(dna_fragment)))$r.squared)


optimal_slope=-1/log10(2)
optimal_E_L=90
optimal_E_H=110

min_R2=0.98


```

### Plotting

```{r}
standard_curves<-ggplot(qPCR_standard,aes(log10(dna_fragment),Ct))+
  geom_point()+
  geom_smooth(method = 'lm')+
  facet_wrap(vars(replicate)) +
  labs(title= "Standard curves per replicate",
       x=expression(log[10]~"[DNA] (fragments/"*mu*"L)",
       y= "Ct (cycles)"))+
  theme_bw()

slope_plot<-ggplot(qPCR_standard_metrics,aes(replicate,slope))+
  geom_point()+
  geom_hline(yintercept =  optimal_slope, col="red")+
  labs(title="Standard curve slope per replicate",
       subtitle = "Ideal slope in red",
       x="Replicate",
       y=expression("Slope"))+
  theme_bw()

E_plot<-ggplot(qPCR_standard_metrics,aes(replicate,E))+
  geom_point()+
  geom_hline(yintercept =  optimal_E_L,col="red")+
  geom_hline(yintercept =  optimal_E_H,col="red")+
  labs(title="Efficiency per replicate",
       subtitle = "Ideal range in red",
       x="Replicate",
       y="Efficiency (%)")+
  theme_bw()

R_squared_plot<-ggplot(qPCR_standard_metrics,aes(replicate,R2))+
  geom_point()+
  geom_hline(yintercept =  min_R2, col="red")+
  labs(title=expression(R^2~" per replicate"),
       subtitle = expression("Minimal "* R^2* " in red"),
       x="Replicate")+
  theme_bw()

standard_curves

slope_plot

E_plot

R_squared_plot

# I would suggest using curve 1, as it performs the best with slope and efficiency, and the Rsquared is sufficiently high. I compute the linear regression 

dna_regression<-lm(log10(dna_fragment)~Ct,qPCR_standard|>filter(replicate=="1"))
```

### Computation

```{r}
qPCR_DNA<-qPCR|>
  filter(sample_class=="sample")|>
  mutate(log_10_dna_frag=predict(dna_regression, newdata = data.frame(Ct = Ct)),
         dna_frag_ul=10^log_10_dna_frag*dilution_factor*5*2)|> #times 5 due to dilution in mastermix, times 2 due to ABB being in 1:2 solution 
  left_join(Design)

Ploidy<-qPCR_DNA|>
  select(channel, dna_frag_ul, replicate )|>
  mutate(channel=factor(channel))|>
  left_join(Broken_cell_count)|>
  mutate(broken_cell_ul = Broken_cell_count_ml/1000)|>
  mutate(ploidy=dna_frag_ul/broken_cell_ul)

Ploidy_per_channel<-Ploidy|>
  group_by(channel)|>
  summarise(ploidy=mean(ploidy),
            sd_ploidy=sd(ploidy))|>
  left_join(Broken_cell_count|>select(channel, phosphate_mM))

Ploidy_per_group<-Ploidy|>
  group_by(phosphate_mM)|>
  summarise(mean_ploidy=mean(ploidy),
            CI1=t.test(ploidy)$conf.int[1],
            CI2=t.test(ploidy)$conf.int[2],
            )

### Since channel 1 had relatively poor lysis efficiency and is an outlier of ploidy in the treatment group, one might consider leaving it out

Ploidy_per_group_no1<-Ploidy|>
  filter(channel!="1")|>
  group_by(phosphate_mM)|>
  summarise(mean_ploidy=mean(ploidy),
            CI1=t.test(ploidy)$conf.int[1],
            CI2=t.test(ploidy)$conf.int[2])



```

### Plotting

```{r}


ggplot(qPCR_DNA,aes(x=factor(channel),y=dna_frag_ul,fill=phosphate_mM))+geom_boxplot(outliers = F, alpha = 0.5)+
geom_jitter(aes(color=factor(dilution_factor)),width = 0.3) +
  scale_fill_brewer(type="seq",palette = "Dark2",name = "[Phosphate] (mM)") +
  scale_color_discrete(palette=c("blue","purple","red"), name="Dilution factor")+
  labs(title = "DNA concentration per channel",
       x="Channel",
       y=expression("[DNA] (fragments/"*mu*"L)"))

ggplot(Ploidy,aes(x=factor(channel),y=ploidy,fill=phosphate_mM))+
  geom_violin()+geom_jitter()+
  scale_fill_brewer(type="seq",palette = "Dark2",name = "[Phosphate] (mM)")+
  labs(title="Violin plot",
       subtitle = "Ploidy per channel",
    x= "[Phosphate] (mM)",
    y="Ploidy (#)"
       ) # Note much more variance in the high phosphate group, as well as in channel 1

ggplot(Ploidy_per_channel, aes(x=channel,y=ploidy,fill=phosphate_mM))+geom_col()+
  labs(title = "Ploidy per channel", 
       x="Channel", y="Ploidy (#)")+
  scale_fill_brewer(type="seq",palette = "Dark2",name = "[Phosphate] (mM)") 

ggplot(Ploidy_per_group, aes(x=phosphate_mM,y=mean_ploidy,fill=phosphate_mM))+
  geom_col(show.legend = F)+
  geom_errorbar(aes(ymin=CI1,ymax=CI2),width=0.2)+
  labs(title = "Ploidy per treatment group",
       subtitle = "Error bars represent a 95% confidence interval",
       x= "[Phosphate] (mM)",
    y="Ploidy (#)" )+
  scale_fill_brewer(type="seq",palette = "Dark2",name = "[Phosphate] (mM)") 

## Since channel 1 had relatively poor lysis efficiency and is an outlier of ploidy in the treatment group, one might consider leaving it out

ggplot(Ploidy_per_group_no1, aes(x=phosphate_mM,y=mean_ploidy,fill=phosphate_mM))+
  geom_col(show.legend = F)+
  geom_errorbar(aes(ymin=CI1,ymax=CI2),width=0.2)+
  labs(title = "Ploidy per treatment group",
       subtitle= "Channel 1 omitted from analysis",
    x= "[Phosphate] (mM)",
    y="Ploidy (#)"
       )+
  scale_fill_brewer(type="seq",palette = "Dark2",name = "[Phosphate] (mM)") # Much tighter CI on the treatment group (makes sense because less variance in the treatment group without 1)

```

### Hypothesis tests

```{r}
# RELATION PLOIDY v EFFICIENCY
summary(lm(ploidy~lysis_efficiency,Ploidy)) # No significant relationship between lysis efficiency and ploidy -> argument in favour of keeping channel 1

round(t.test(ploidy~phosphate_mM,Ploidy)$p.value,3)

t.test(ploidy~phosphate_mM,Ploidy|>filter(channel!="1"))

# Normality of ploidy
Ploidy|> group_by(phosphate_mM)|>
  summarise(n=n(),
            "p-value"=round(shapiro.test(ploidy)$p.value ,3))|>
  mutate("[Phosphate] (mM)"=phosphate_mM)|> 
  select(-phosphate_mM)|>
  flextable(col_keys = c("[Phosphate] (mM)", "n", "p-value"))

# Ploidy is not normally distributed for treatment group, due to channel 1 being outlier?

Ploidy|> filter(channel!=1)|>
  group_by(phosphate_mM)|>
  summarise(n=n(),
            "p-value"=round(shapiro.test(ploidy)$p.value ,3))|>
  mutate("[Phosphate] (mM)"=phosphate_mM)|>
  select(-phosphate_mM)|>
  flextable(col_keys = c("[Phosphate] (mM)", "n", "p-value"))# Without channel 1 normality is no longer ruled out by shapiro wilks test -> argument against keeping channel 1


```

### Additional analysis

#### Does the OD_720 indeed fit linearly with the cell count?

From the data below it seems to indeed be the case, that's nice!

```{r}
OD_v_CellCount<-MC_CASY_cellcount|>
  left_join(SpecPhot|>select(channel,od_value_730,od_value_720)|>
              mutate(channel=factor(channel)))
# Plot
ggplot(OD_v_CellCount, aes(x = mean_count_ml)) +
  geom_point(aes(y = od_value_730, colour = "OD 730")) +
  geom_smooth(aes(y = od_value_730, colour = "OD 730"), method = "lm") +
  geom_point(aes(y = od_value_720, colour = "OD 720")) +
  geom_smooth(aes(y = od_value_720, colour = "OD 720"), method = "lm") +
  scale_colour_manual(
    name = "Measurement",
    values = c("OD 730" = "blue", "OD 720" = "red")
  )+
  labs(title="OD versus mean cell count",
       x="Mean cell count (#/ml)",
       y=expression(OD))

# Run tests
summary(lm(od_value_730~mean_count_ml,OD_v_CellCount)) 
summary(lm(od_value_720~mean_count_ml,OD_v_CellCount))

# Fairly low R^2 probably due to small sample size 
```

### Cell size vs treatment group

Assuming sphericality of cells the CASY counter gives a read of the cell
size. We compare these measurements to the treatment group for the BBB
samples, as these will also be used in further analysis comparing the
cell size with ploidy.

The plots below suggest that both mean and peak cell diameter go down in
the control group with more phosphate. I found a Cohen's d value of 4.82
(CI\[2.90, 6.69\]), suggesting a statistically significant but small
effect.

This feels surprising to me, and I have so far not found anything
concrete in the literature on this. It could be possible that the cells
grow larger before dividing, leading to larger but fewer cells in the
phosphate-starved population. As shown in the spectrophotometry data,
there is a decrease in $\text{OD}_{730}$ in the treatment group.

```{r}
size_v_treatment<-CASY|>
  filter(sample_type=="BBB")|>
  select(channel,mean_diameter,peak_diameter)|>
  left_join(Design,by="channel")

summary(lm(data=size_v_treatment,formula=mean_diameter~phosphate_mM))

ggplot(size_v_treatment,aes(x=factor(phosphate_mM),
                            y=mean_diameter,
                            fill=factor(phosphate_mM)))+
  geom_boxplot(show.legend = F)+geom_jitter(show.legend = F,width = 0.1)+
  scale_fill_brewer(type="seq",palette = "Dark2",name = "[Phosphate] (mM)")+
  labs(title = "Mean cell diameter per treatment group",
    x= "[Phosphate] (mM)",
    y= expression("Mean diameter ("*mu*"m)")
       )

ggplot(size_v_treatment,aes(x=factor(phosphate_mM),
                            y=peak_diameter,
                            fill=factor(phosphate_mM)))+
  geom_boxplot(show.legend = F)+geom_jitter(show.legend = F,width = 0.1)+
  scale_fill_brewer(type="seq",palette = "Dark2",name = "[Phosphate] (mM)")+
  labs(title = "Peak cell diameter per treatment group",
    x= "[Phosphate] (mM)",
    y= expression("Peak diameter ("*mu*"m)")
       )

library(effectsize)

cohens_d(mean_diameter~phosphate_mM,data=size_v_treatment)


```

# Omics data analysis

### Load data

```{r}
Transcriptomics<-read_csv("Data/log2FC_transcriptomics.csv")

Proteomics<-read_tsv("Data/Proteomics_Metabolomics/proteinGroups.txt")

Metabolomics<-read_excel("Data/Proteomics_Metabolomics/HILIC_BA_Metabolites_Synechocystis.xlsx")

Lipidomics<-read_excel("Data/Proteomics_Metabolomics/C18_CD_Lipids_SILS54_Synechocystis.xlsx")

Metadata<-read_excel("Data/Proteomics_Metabolomics/MetaData.xlsx")
```

## Transcriptomics analysis

## Multiomics Wrangling

```{r}
vars<-c("ID","WT-S1","WT-S2","WT-S3","WT-S4", "Treatment-S5","Treatment-S6","Treatment-S7","Treatment-S8")

split_cols <- strsplit(vars[-1], "-")

header2 <- c("Group", sapply(split_cols, `[`, 1))
header1 <- c("Name",   sapply(split_cols, `[`, 2))

# Proteomics
Proteomics_fixed<-Proteomics|>
  select(contains("LFQ"), `Majority protein IDs`, `Only identified by site`, Reverse, `Potential contaminant`)|>
  filter(is.na(`Only identified by site`),
         is.na(Reverse),
         is.na(`Potential contaminant`))|>
  select(`Majority protein IDs`,contains("LFQ"))

colnames(Proteomics_fixed)<-vars

# write to file
proteomics_file <- file("Omics_Data_Output/Proteomics.csv", "w")

writeLines(paste(header1, collapse = ","), proteomics_file)
writeLines(paste(header2, collapse = ","), proteomics_file)

# then append your data (without colnames)
write.table(Proteomics_fixed, proteomics_file, sep = ",", row.names = FALSE, col.names = FALSE)

# Metabolomics
Metabolomics_fixed<- Metabolomics|>
  unite("ID",c(`RT [min]`,`m/z meas.`))|>
  mutate(ID = paste0(ID, "_polar"))|>
  select(ID,13:20)

colnames(Metabolomics_fixed)<-vars

# write to file
metabolomics_file <- file("Omics_Data_Output/Metabolomics.csv", "w")

writeLines(paste(header1, collapse = ","), metabolomics_file)
writeLines(paste(header2, collapse = ","), metabolomics_file)

# then append your data (without colnames)
write.table(Metabolomics_fixed, metabolomics_file, sep = ",", row.names = FALSE, col.names = FALSE)


Lipidomics_fixed<- Lipidomics|>
  unite("ID",c(`RT [min]`,`m/z meas.`))|>
  mutate(ID = paste0(ID, "_apolar"))|>
  select(ID,13:20)

colnames(Lipidomics_fixed)<-vars

# write to file
lipidomics_file <- file("Omics_Data_Output/lipidomics.csv", "w")

writeLines(paste(header1, collapse = ","), lipidomics_file)
writeLines(paste(header2, collapse = ","), lipidomics_file)

# then append your data (without colnames)
write.table(Lipidomics_fixed, lipidomics_file, sep = ",", row.names = FALSE, col.names = FALSE)


multi_omics<-rbind(Proteomics_fixed,Metabolomics_fixed,Lipidomics_fixed)|>
  distinct(ID,.keep_all = T)



# write to file
multi_file <- file("Omics_Data_Output/Multiomics.csv", "w")

writeLines(paste(header1, collapse = ","), multi_file)
writeLines(paste(header2, collapse = ","), multi_file)

# then append your data (without colnames)
write.table(multi_omics, multi_file, sep = ",", row.names = FALSE, col.names = FALSE)

```

## Transcriptomics

```{r}
threshold_FC<-2
threshold_pvalue<-0.005

points_to_colo_red<-Transcriptomics|>
  filter(log2FoldChange>threshold_FC,
         padj<threshold_pvalue)

points_to_colo_blue<-Transcriptomics|>
  filter(log2FoldChange< -threshold_FC,
         padj<threshold_pvalue)

genes_upregulated<-points_to_colo_red$gene_name
genes_downregulated<-points_to_colo_blue$gene_name

ggplot(Transcriptomics,aes(x=log2FoldChange,y=-log10(padj)))+geom_point()+
  geom_vline(xintercept =  threshold_FC,col="red")+
  geom_vline(xintercept =  -threshold_FC,col="red")+
  geom_hline(yintercept = -log10(threshold_pvalue),col="blue")+geom_point(
  data = points_to_colo_red,
  aes(x = log2FoldChange, y = -log10(padj)),
  color = "red")+
  geom_point(data = points_to_colo_blue,
  aes(x = log2FoldChange, y = -log10(padj)),
  color = "blue")

write(genes_upregulated,"upreg_trans.csv")


write(genes_downregulated,"downreg_trans.csv")
```
