---
title: "Synechocystis_Data_Analysis"
author: "T.A.J. Uijen, "
date: "01/04/2025-29/05/2026"
output:
  pdf_document:
    number_sections: no
    toc: yes
    toc_depth: 3
  html_document:
    toc: yes
    toc_depth: '3'
    df_print: paged
editor_options:
  markdown:
    wrap: 72
execute: 
  echo: true
  message: false
  warning: false
---

# Importing data and loading packages

```{r, echo=TRUE, message=FALSE}
library(readr)
library(tidyr)
library(dplyr)
library(ggplot2)
library(flextable)
library(GGally)
library(broom)

Design<-read_csv("Data/Design.csv")
SpecPhot<-read_csv("Data/Spectrophotometer.csv")
Turbidostat<-read_csv("Data/Turbidostat.csv")
CASY<-read_csv("Data/Casy.csv")
qPCR<-read_csv("Data/Qpcr.csv")

```

# Analysis of turbidostat and OD data

## Spectrophotometer data

### A bit of data wrangling

```{r}
Design<-Design|>
  mutate(phosphate_mM=factor(phosphate_mM))

SpecPhot_total<-SpecPhot|>
  arrange(channel)|>
  mutate(phosphate_mM=Design$phosphate_mM,
         chlor_a=10.186*(od_value_685-od_value_730)-0.08,
         PBS=(od_value_634-od_value_730)/od_value_730)
  
# Compute the mean for each treatment group
SpecPhot_treatment<-SpecPhot_total|>
  group_by(phosphate_mM)|>
  summarise(od_value_730=mean(od_value_730),
            od_value_685=mean(od_value_685),
            od_value_634=mean(od_value_634),
            od_value_720=mean(od_value_720),
            chlor_a=mean(chlor_a),
            PBS=mean(PBS))


```

### Visualising the data

```{r}
od_730_plot<-ggplot(SpecPhot_treatment, aes(phosphate_mM,od_value_730,fill = phosphate_mM))+
  geom_col(show.legend = F)+
  labs(title="Bar plot",
       subtitle = expression(OD[730]~"per treatment group"),
       x="[Phosphate] (mM)", 
       y=expression(OD[730]))+
  scale_fill_brewer(type="seq",palette = "Dark2") +
  theme_bw()


chlor_a_plot<-ggplot(SpecPhot_treatment, aes(phosphate_mM,chlor_a,fill = phosphate_mM))+geom_col(show.legend = F)+
  labs(x="[Phosphate] (mM)", 
       y=expression("Chlorophyl A (mg/L)"),
       title="Bar plot",
       subtitle = "Chorophyl A per treatment group")+
  scale_fill_brewer(type="seq",palette = "Dark2")+
  theme_bw()

pbs_plot<-ggplot(SpecPhot_treatment, aes(phosphate_mM,PBS,fill = phosphate_mM))+
  geom_col(show.legend = F)+
  labs(x="[Phosphate] (mM)", 
       y=expression("Phycobilisomes (mg/L)"),
       title="Bar plot",
       subtitle = "Phycobilisomes per treatment group")+
  scale_fill_brewer(type="seq",palette = "Dark2")+
  theme_bw()

od_730_plot
chlor_a_plot
pbs_plot
```

Observations:

-   The OD is higher in the group with more phosphate

-   The Chlorophil A content is higher in the group with more phosphate.
    This is noteworthy since there is less light provided to the
    cyanobacteria here. One might propose that the bacteria attempt to
    get more out of the limited light?

-   Phycobilisome content does not change significantly, this is also
    shown with a t-test below

### Hypothesis tests

```{r}
test_OD<-t.test(od_value_730~phosphate_mM,SpecPhot_total)
test_chlorA<-t.test(chlor_a~phosphate_mM,SpecPhot_total)
test_PBS<-t.test(PBS~phosphate_mM,SpecPhot_total)

spec_phot_test_results<-data.frame(Variable=c('OD_730',"Chlorophil A", "Phycobilisomes"), P_value=c(round(test_OD$p.value,3),  round(test_chlorA$p.value,3), round(test_PBS$p.value,3)))

spec_phot_test_results|>flextable()
```

## Turbidostat data

### Initial visualisation of time series

```{r}
Turbidostat<-Turbidostat|>
  left_join(Design|>select(channel,phosphate_mM))

ggplot(Turbidostat,aes(time_h,OD,colour=phosphate_mM))+
  geom_point()+
  facet_wrap(~channel,scales="free_x")+
  labs(title="Time series",subtitle=expression(OD[720] ~" over time in hours, coloured by treatment group"), x="Time (h)",y=expression(OD[720]), colour="[phosphate] (mM)")+
  scale_colour_brewer(type="seq",palette = "Dark2")+
  theme_bw()
  
```

Observations:\

-   Channel 5 starts low, behaves normally after the first activation of
    the pump

-   Channel 3 seems translated downwards, as the pump is supposed to
    turn on at OD=0.35, here this seems to happen at OD=0.32. In
    principle this does not change the analysis, as we are concerned
    with the rate of change, not the absolute values.

### Some data wrangling again

```{r}
Growth_rates<-Turbidostat|>
  mutate(log_OD=log(OD))|> # compute log OD
  group_by(channel,decision)|>
  #Compute growth rates and R^2
  summarise(mu=coef(lm(log_OD~time_h))[2],
            r_sq=summary(lm(log_OD~time_h))$r.squared, 
            .groups = "drop_last")|>  
  filter(r_sq>.9) |># Remove cycles with less than 0.9 R^2|. This causes ch. 1 to only have 1 point
  left_join(Design |> select(channel, phosphate_mM))


test<-shapiro.test(Growth_rates|>filter(phosphate_mM=="0.0115")|>pull(mu))
Mean_growth<-Growth_rates|>
  group_by(channel)|>
  summarise(mean_mu=mean(mu),
            sd_mu=sd(mu))|>
  left_join(Design |> select(channel, phosphate_mM), by = "channel")
  

Growth_rates<-Growth_rates|>
  mutate(channel=factor(channel))

Mean_growth_per_group<-Growth_rates|>
  group_by(phosphate_mM)|>
  summarise(mean_mu_grp=mean(mu),sd_mu_grp=sd(mu),
            CI1=t.test(mu)$conf.int[1],
            CI2=t.test(mu)$conf.int[2]) # Confidence interval is assuming normal distribution of the mean




```

### Visualisation

```{r}
violin_plot<-ggplot(Growth_rates,aes(channel,mu,fill=phosphate_mM))+
  geom_violin()+
  geom_point()+
  labs(title="Violin plot",
       subtitle = "Growth rates per channel, coloured by treatment group",
       x="Channel",
       y=expression(mu),
       fill="[phosphate] (mM)")+
  scale_fill_brewer(type="seq",palette = "Dark2")+
  theme_bw()

technical_var<-ggplot(Mean_growth,aes(channel,mean_mu,fill=phosphate_mM))+
  geom_col()+
  geom_errorbar(aes(ymin=mean_mu-2*sd_mu,ymax=mean_mu+2*sd_mu),width=0.2)+
  scale_fill_brewer(type="seq",palette = "Dark2")+
  labs(title="Bar plot",
       subtitle = "Mean growth rates per channel, error bars represent 2 standard deviations",
       x="Channel",
       y=expression("mean" ~ mu),
       fill="[phosphate] (mM)")+
  theme_bw()

biological_var<-ggplot(Mean_growth_per_group,aes(phosphate_mM,mean_mu_grp,fill=phosphate_mM))+geom_col(show.legend = F)+geom_errorbar(aes(ymin=CI1,ymax=CI2),width=0.2)+
  scale_fill_brewer(type="seq",palette = "Dark2")+
  labs(title="Bar plot",
       subtitle = "Growth rates per treatment group, error bars represent 95% confidence interval",
       x="[phosphate] (mM)",
       y=expression("mean"~ mu))+
  theme_bw()
violin_plot

technical_var

biological_var

t.test(mu~phosphate_mM,Growth_rates) #the null hypothesis that the mean growthrate between groups is the same is not rejected 
```

Observations:

-   No measurement of SD of channel 1

-   The other channels seem to not differ significantly (assuming normal
    distributions)

-   There seems to be no group difference between phosphate
    concentration (this is as desired) This is corroborated by a t-test
    (under the assumption of normality.

We test the normality assumption using a Shapiro-Wilk normality test,
given the small sample size of 20 growth rates that have an $R^2>0.90$

```{r}
normality_test<-Growth_rates|>
  group_by(phosphate_mM)|>
  summarise(p.value=shapiro.test(mu)$p.value)
```

The test does not suggest non-normality of the data, which speaks in
favour of the reliability of the analysis above.

Since the tests do not suggest non-normality, nor a difference in mean
OD$_{720}$, we can test normality further by not grouping by treatment:

```{r}
shapiro.test(Growth_rates$mu)
```

In this case we also do not reject the null hypothesis of normality.

# Ploidy determination

## CASY-counter data

Assumptions: Cells are spherical, no contaminants not captured in the
blank.

Counts measured by the CASY-counter are in a sample that is undiluted
for the MC samples, 1:50 for the BBB samples and 1:2 for the ABB
samples. 10 $\mu$L of this dilution is then added to 10ml of CASY
buffer. Of this mixture, 200 $\mu$L is measured by the CASY-counter.

First compute the mean cell counts for each channel over the 3
replicates

```{r}
MC_CASY_cellcount<-CASY|>
  filter(sample_type=="MC")|>
  group_by(channel)|>
  summarise(mean_count_ml=mean(1001*5*counts))|> # Multiply by 5 for counts per ml
  left_join(Design)|>
  mutate(channel=factor(channel)) # makes the plots prettier

BBB_cellcount<-CASY|>
  filter(sample_type == "BBB")|>
  group_by(channel)|>
  summarise(BBB_count_ml=mean(1001*5*50*counts))|>
  left_join(Design) # Multiply by 50 to account for dilution

ABB_BBB_cellcount<-CASY|>
  filter(sample_type=="ABB")|>
  group_by(channel)|>
  summarise(ABB_count_ml=mean(1001*5*2*counts))|> # Multiply by 2 to account for dilution
  left_join(BBB_cellcount) 

Broken_cell_count=ABB_BBB_cellcount|>
  mutate(Broken_cell_count_ml=BBB_count_ml-ABB_count_ml,
         channel=factor(channel),
         lysis_efficiency = 100*Broken_cell_count_ml/BBB_count_ml)
  
ABB_BBB_cellcount<-ABB_BBB_cellcount|>
  pivot_longer(cols = c(ABB_count_ml,BBB_count_ml),
               names_to = "Bead_status",
               values_to = "count_ml") |>
  mutate(channel=factor(channel),
         Bead_status=factor(Bead_status))
 
  



```

```{r}
MC_cellcount_plot<-ggplot(MC_CASY_cellcount,aes(channel,mean_count_ml,fill=phosphate_mM))+geom_col()+
  scale_fill_brewer(type="seq",palette = "Dark2",name = "[Phosphate] (mM)") +
  theme_bw()+
  labs(title="Cell counts in the MultiCultivator per channel",
       subtitle = "Computed from CASY-counts",
       x="Channel",
       y= "Cell concentration (cells/ml)")


comp_cellcount_plot<-ggplot(ABB_BBB_cellcount,aes(channel,count_ml,fill=Bead_status))+
  geom_col(position = "dodge") +
  scale_fill_brewer(type="seq",palette = "Set1",name=" Bead beating", labels=c('ABB', "BBB")) +
  labs(title="Cell counts before and after bead beating per channel",
       subtitle = "Computed from CASY-counts",
       x="Channel",
       y= "Cell concentration (cells/ml)") + 
  theme_bw()

Broken_cell_count_plot<-ggplot(Broken_cell_count,aes(channel,Broken_cell_count_ml,fill=phosphate_mM))+geom_col()+
  scale_fill_brewer(type="seq",palette = "Dark2",name = "[Phosphate] (mM)") +
  labs(title="Broken cell counts per channel",
       subtitle = "Computed from CASY-counts",
       x="Channel",
       y= "Cell concentration (cells/ml)") +
  theme_bw()

lysis_efficiency_plot<-ggplot(Broken_cell_count,aes(channel,lysis_efficiency,fill=phosphate_mM))+geom_col()+
  scale_fill_brewer(type="seq",palette = "Dark2",name = "[Phosphate] (mM)") +
  labs(title="Lysis efficiency per channel",
       subtitle = "Computed from CASY-counts",
       x="Channel",
       y= "Lysis efficiency (%)") +
  theme_bw()

MC_cellcount_plot

comp_cellcount_plot

Broken_cell_count_plot
lysis_efficiency_plot
```

## qPCR Data:
